Effective Inclusion of Electronic Polarization to a Negatively Charged Membrane

### 2019-03-08 Manuscript/manuscript.tex
Update text

\subsection{Validation of membrane properties}

Cation binding affinity to phosphatidylcholine (PC) lipid bilayers can be evaluated using 
the order parameters of $\alpha$ and $\beta$ carbons in the lipid headgroup 
(see Fig.~\ref{fig:chemstruct_pc_ps} for the labeling) \cite{catte16}.
According to the ''electrometer concept'' \citep{seelig87}, these order parameters decrease
proportionally to the bound charge and can therefore be used to quantify the cation binding
affinity to lipid bilayers. The situation is more complicated for lipid bilayers containing
PS lipids, for which there is no ion-free reference state due to the presence of counterions.
Moreover, additional ions often lead to phase transitions in such bilayers \cite{nmrlipids_proj4,feigenson86,mattai89,roux91,roux90}
Here, we overcome this complication by comparing simulations to the experimental data 
measured from POPC:POPS (5:1) mixtures, 
which are not as highly negatively charged, 
and do not ondergo phase transitions with additional ions \cite{roux90} 
as was also demonstrated in the recent work by the NMRlipids project \cite{nmrlipids_proj4}.

%experimentally quantified
%by measuring the changes of the head group order parameters in the lipids using solid state NMR.
%In the case of 
%such changes are known under the term 
%It is experimentally observed that the \ce{C-H} bond
%order parameters of $S^\alpha$ and $S^\beta$ carbons in a PC lipid head group
%are proportional to the amount of charge, positive or negative, bound per lipid~\citep{seelig87, roux90}.
%The change of the order parameters measured with varying concentrations of aqueous ions 
%can be then related to the amount of bound ions.

@@

\begin{figure}[tbp!] 
  \centering 
  \includegraphics[height=\figheight]{../img/ecc_pops/density_profiles_na-k-counterions_wat_phos_compar_purePOPS_ecclipids-lipid17.pdf}
  \includegraphics[height=\figheight]{../img/ecc_pops/density_profiles_na-k-counterions_wat_phos_compar_5PC-1PS_ecclipids-lipid17.pdf}
  \caption{\label{fig:POPS-counterions-dens}
    Number density profiles of \ce{K^{+}}, \ce{Na^{+}} counterions along the membrane normal axis 
    for the negatively charged bilayers composed of only POPS (left) or POPC:POPS (5:1, right) 
    using ECC-lipids/ECC-ions (solid lines) and Lipid17/Dang ions (dashed lines).  
    The density profiles of phosphate groups and water are divided by 4 and 100, respectively.  
}
\todo{JOE: add labels with the relative surface excess $\Gamma$, which are promised to be present in the text.}
\todo{JOE: split this figure into two: separate for POPS (to SI) and PC:PS 5:1 (main text)?}
\end{figure}

\begin{figure}[tb!] 
  \centering 
  \includegraphics[width=\figwidth]{../Fig/lipids_chemfig_POPC_POPS.pdf} 
  \caption{ \label{fig:chemstruct_pc_ps} 
            The chemical structure of POPC and POPS 
            including the labeling of the carbon segments
            used in the measurements of order parameteres. 
  }  
\end{figure}

@@

\subsection{Binding of counterions to negatively charged PC:PS bilayers and interactions between PC and PS head groups}

According to the electrometer concept,
the head group order parameters $\alpha$ and $\beta$ of PC lipids
change proportionally to the amount of bound charge
in a fixed ratio. \cite{seelig87,scherer87,roux90} 
This was demonstrated for a bound positive charge in simulations with POPC and 
cationic surfactants or aquaeous cations employing the ECC-POPC model \cite{melcr18}.
In this work, we will use the electrometer concept also for a bound negative charge 
using PS lipids as negatively charged surfactants. \cite{roux90}

@@

@@

\subsection{Molecular interaction and binding affinities of K$^+$ and Na$^+$ cations to mixed POPC:POPS (5:1) membrane} 
\label{sec:K_Na_added}

@@

% This§: 
% NA/K binding status, 
% how we tried finding a model (NMRlipids4), and 
% how it was futile
% 
The binding of \ce{Na^+} cations to phospholipid bilayers 
is not generally agreed on between simulations 
\citep{bockmann03,sachs04,berkowitz06,cordomi09}
and experiments 
\citep{cevc90,tocanne90,hauser76,herbette84,uhrikova08}.
We attempted to find a model, 
which would successfully interpret the experiments on binding of \ce{Na^+}
to neutral and negatively charged phospholipid bilayers \citep{akutsu81, roux90} 
in our works \citep{catte16, nmrlipids_proj4}.
After testing the currently available force fields for phospholipids,
we concluded
that the interactions are generally overestimated in magnitude in almost all models 
but Lipid14 (PC) \citep{dickson14}, resp. Lipid17 (PS) \citep{lipid17-future}. 
This model yields a semi-quantitative agreement with the experimentally measured small changes of the order parameters \cite{catte16}
when used with the model of ions by \citet{aqvist90}. 
However, when used with a more accurate model of ions by \citet{Pluharova2014, martinek17},
the model overestimates the binding affinity of \ce{Na^+}
measured with lipid electrometer concept. \citep{melcr18}
In total, these results suggest that improvements 
in the lipid parameters are required for more accurate interactions even with monovalent cations. 
\citep{catte16, melcr18, nmrlipids_proj4}

% This§: 
% we have developed a model that can describe the NMR exp. using ECC
% 
Generally improved behaviour 
of the POPC and POPS head group order parameters 
with \ce{NaCl} or \ce{KCl} concentrations 
was achieved through the combination of models 
ECC-lipids and ECC-ions \citep{martinek17, kohagen16, Pluharova2014}. 
Simulations with these models 
reveal a good agreement with the NMR experiments 
for both neutral and negatively charged membranes.
The results are plotted for \ce{NaCl} in Fig. \ref{fig:delta_ordPar_NaCl_PCPS}, 
and for \ce{KCl} in Fig. \ref{fig:delta_ordPar_KCl_PCPS}.

>>>>

%\subsection{Validation of membrane properties}
% SAMULI: I have dissolved this section into results and discussion

@@

@@

\subsection{Counterion binding affinity and interactions between PC and PS head groups}

Binding affinity of monovalent ions to lipid bilayers depends strongly on
force field parameters \cite{catte16,melcr18,nmrlipids_proj4},
which is the case also here as the sodium binding affinity to
POPS lipid bilayer is stronger in Lipid17 simulation with Dang ions than in
ECC-POPS simulation (Fig. \ref{fig:POPS-counterions-dens}).
The different cation binding affinities to lipid bilayers have been previously evaluated
against experiments using the order parameters of $\alpha$ and $\beta$ carbons in the
phosphatidylcholine (PC) lipid headgroup (see Fig.~\ref{fig:chemstruct_pc_ps} for the labeling) \cite{catte16,melcr18,nmrlipids_proj4}.
According to the ''electrometer concept'' \citep{seelig87}, these order parameters decrease
proportionally to the bound charge and can therefore be used to quantify the cation binding
affinity to lipid bilayers. Because the response of PS lipid headgroup order parameters to the bound charge
is less well undestood and addition of cation may cause phase transitions in negatively charged
bilayers \cite{feigenson86,mattai89,roux91,roux90}, the cation binding affinity to bilayers with
PS lipids has evaluated measuring the PC headgroup order parameters from POPC:POPS (5:1)
mixtures\cite{roux90,nmrlipids_proj4}.

The situation is more complicated for lipid bilayers containing
PS lipids, for which there is no ion-free reference state due to the presence of counterions.
Moreover, additional ions often lead to phase transitions in such bilayers \cite{nmrlipids_proj4,feigenson86,mattai89,roux91,roux90}

@@

%According to the electrometer concept,
%the head group order parameters $\alpha$ and $\beta$ of PC lipids
%change proportionally to the amount of bound charge
%in a fixed ratio. \cite{seelig87,scherer87,roux90} 
%This was demonstrated for a bound positive charge in simulations with POPC and 
%cationic surfactants or aquaeous cations employing the ECC-POPC model \cite{melcr18}.
%In this work, we will use the electrometer concept also for a bound negative charge 
%using PS lipids as negatively charged surfactants. \cite{roux90}

%is the
%is not generally agreed on between simulations 
%\citep{bockmann03,sachs04,berkowitz06,cordomi09}
%and experiments 
%\citep{cevc90,tocanne90,hauser76,herbette84,uhrikova08}.
%We attempted to find a model, 
%which would successfully interpret the experiments on binding of \ce{Na^+}
%to neutral and negatively charged phospholipid bilayers \citep{akutsu81, roux90} 
%in our works \citep{catte16, nmrlipids_proj4}.
%After testing the currently available force fields for phospholipids,
%we concluded
%that the interactions are generally overestimated in magnitude in almost all models 
%but Lipid14 (PC) \citep{dickson14}, resp. Lipid17 (PS) \citep{lipid17-future}. 
%This model yields a semi-quantitative agreement with the experimentally measured small changes of the order parameters \cite{catte16}
%when used with the model of ions by \citet{aqvist90}. 
%However, when used with a more accurate model of ions by \citet{Pluharova2014, martinek17},
%the model overestimates the binding affinity of \ce{Na^+}
%measured with lipid electrometer concept. \citep{melcr18}
%In total, these results suggest that improvements 
%in the lipid parameters are required for more accurate interactions even with monovalent cations. 
%\citep{catte16, melcr18, nmrlipids_proj4}

\begin{figure}[tbp!] 
  \centering 
  \includegraphics[height=\figheight]{../img/ecc_pops/density_profiles_na-k-counterions_wat_phos_compar_purePOPS_ecclipids-lipid17.pdf}
  \includegraphics[height=\figheight]{../img/ecc_pops/density_profiles_na-k-counterions_wat_phos_compar_5PC-1PS_ecclipids-lipid17.pdf}
  \caption{\label{fig:POPS-counterions-dens}
    Number density profiles of \ce{K^{+}}, \ce{Na^{+}} counterions along the membrane normal axis 
    for the negatively charged bilayers composed of only POPS (left) or POPC:POPS (5:1, right) 
    using ECC-lipids/ECC-ions (solid lines) and Lipid17/Dang ions (dashed lines).  
    The density profiles of phosphate groups and water are divided by 4 and 100, respectively.  
}
\todo{JOE: add labels with the relative surface excess $\Gamma$, which are promised to be present in the text.} \\
\todo{JOE: split this figure into two: separate for POPS (to SI) and PC:PS 5:1 (main text)? SAMULI: Should be POPS in main text, mixture in SI.}
\end{figure}

\begin{figure}[tb!] 
  \centering 
  \includegraphics[width=\figwidth]{../Fig/lipids_chemfig_POPC_POPS.pdf} 
  \caption{ \label{fig:chemstruct_pc_ps} 
            The chemical structure of POPC and POPS 
            including the labeling of the carbon segments
            used in the measurements of order parameteres. 
          }
          \todo{The POPS structure and labeling is already in figure \ref{simVSexpNOions_POPS}.
            The POPC structure and labeling could be added on top of the new version of figure \ref{fig:delta_ordPar_NaCl_PC-PS_mix} with only PC,
          where PC results occur in the first time, and this figure could be removed.}
\end{figure}

Generally improved behaviour 
of the POPC and POPS head group order parameters 
with \ce{NaCl} or \ce{KCl} concentrations 
was achieved through the combination of models 
ECC-lipids and ECC-ions \citep{martinek17, kohagen16, Pluharova2014}. 
Simulations with these models 
reveal a good agreement with the NMR experiments 
for both neutral and negatively charged membranes.
The results are plotted for \ce{NaCl} in Fig. \ref{fig:delta_ordPar_NaCl_PCPS}, 
and for \ce{KCl} in Fig. \ref{fig:delta_ordPar_KCl_PCPS}.

@@

%\subsection{Molecular interaction and binding affinities of K$^+$ and Na$^+$ cations to mixed POPC:POPS (5:1) membrane} 
%\label{sec:K_Na_added}

@@
